---
title: "Model 3"
author: "Leonard Dwight Soledad"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# Helper Packages And Modeling Packages
```{r}
# Load Helper Packages
library(dplyr)    
library(ggplot2)   
library(stringr)  
library(gridExtra)
library(readr)
library(bestNormalize)

# Load Modeling Packages
library(cluster)    
library(factoextra)
library(tidyverse)
library(mclust)

```

*CLUSTERING TECHNIQUES*

In this model, we will execute and differentiate the clustering techniques results from K-Means, Hierarchical and Model based clustering disregarding the binary output and categorical variables in the data set. We will start by normalizing the data *"radiomics_completedata"*.



#     Loading the Data
#     The data contains 197 rows and 431 columns with *Failure.binary* binary output.
```{r}

Model3Arawdata <- read_csv("radiomics_completedata.csv")
Model3Arawdata
```


#       *Data Reprocessing*


#   Checking for null and missing values
#   We are using *anyNA()* function to determine if there is any missing value in the data.
```{r}

anyNA(Model3Arawdata)

#The output will show either *True* or *False*. There are missing values If True, thus you have to omit the missing values using *na.omit()*. Otherwise, False.
  
#[1] FALSE

# The result is False, hence, the data has no missing values.
```

*Checking the Normality of the Data*
We are using *Shapiro-Wilk's Test* to check the normality of the data.

```{r,warning=F}

Model3Anumrawdata <- Model3Arawdata%>%select_if(is.numeric) 

Model3Anumrawdata <- Model3Anumrawdata[ , -1]

Model3ASWtest <- apply(Model3Anumrawdata, 2, function(x){shapiro.test(x)})
```


Next we need to list only the p-value of the respective variables to proceed with the test. We are using the *unlist()* and *lapply()* functions to achieve this goal.

```{r}

Model3ADRpvalue <- unlist(lapply(Model3ASWtest, function(x) x$p.value))

```


```{r}

sum(Model3ADRpvalue<0.05)  # not normally distributed
sum(Model3ADRpvalue>0.05)  # normally distributed
Model3ASWtest$Entropy_cooc.W.ADC

# [1] 428
# [1] 1

#  Currently, there are 428 variables that are not normally distributed and only the Entropy_cooc.W.ADC is normally distributed.
```

The goal is that all variables should be normally distributed.
Next, we are using *orderNorm()* function. And we need to exclude the *Entropy_cooc.W.ADC* since it is already normally distributed.

```{r,warning=F}
Model3ADRtransrawdata <- Model3Arawdata[,c(3,5:length(names(Model3Arawdata)))]

Model3ADRtransrawdata <- apply(Model3ADRtransrawdata,2,orderNorm)

Model3ADRtransrawdata <- lapply(Model3ADRtransrawdata, function(x) x$x.t)

Model3ADRtransrawdata <- Model3ADRtransrawdata%>%as.data.frame()

Model3ASWtest <- apply(Model3ADRtransrawdata,2,shapiro.test)

Model3ASWtest <- unlist(lapply(Model3ASWtest, function(x) x$p.value))
```

Next, we will be testing the data to check the normality or the transformed data.
```{r,warning=F}
sum(Model3ASWtest <0.05)  # for not normally distributed
sum(Model3ASWtest >0.05)  # for normally distributed

#[1] 0
#[1] 428

# Now, the 428 variables that were initially not normally distributed are now normally distributed.
```


```{r}

Model3Arawdata[,c(3,5:length(names(Model3Arawdata)))]=Model3ADRtransrawdata

```

We are getting the correlation of the whole data except the categorical variables
```{r}

Model3ACorrMat = cor(Model3Arawdata[,-c(1,2)])
heatmap(Model3ACorrMat,Rowv=NA,Colv=NA,scale="none",revC = T)

```

Finally, we will convert the data frame output of data reprocessing into "csv" file, which will we use for the entire Final Project.

```{r}
library(data.table)

fwrite(Model3Arawdata, "Model3_Final_Project_Data.csv")
```

Lastly, let's check if the dataframe we have exported to CSV is really the normal data.
```{r}
Model3Arawdata1 <- read_csv("Model3_Final_Project_Data.csv")
Model3Arawdata1

Model3Anumrawdata1 <- Model3Arawdata1%>%select_if(is.numeric) 

Model3Anumrawdata1 <- Model3Anumrawdata1[ , -1]

Model3ASWtest1 <- apply(Model3Anumrawdata1, 2, function(y){shapiro.test(y)})

Model3ADRpvalue1 <- unlist(lapply(Model3ASWtest1, function(y) y$p.value))

sum(Model3ADRpvalue1<0.05)  # not normally distributed
sum(Model3ADRpvalue1>0.05)  # normally distributed

#Yes! We were able to produce the correct CSV file and we are now ready to use it for the entire project.
```



**MODEL 3**

```{r}

# Load the dataset
model3data <- read_csv("Model3_Final_Project_Data.csv")
model3data
```

# Scaling/Standardizing the Data
```{r}
model3fdata <- scale(model3data[c(3:431)])
head(model3fdata)
sum(is.na(model3fdata))
```


1. K-Means Clustering

K-Means Clustering is one of the most well-known and commonly used clustering algorithms for partitioning observations into a set of k groups. 

The primary objective of k-means clustering is to make clusters within-cluster variation that is minimized. We will perform K-means clustering with 3 clusters, 100 maximum number of iterations, and 100 n start. 


_____***K-MEANS CLUSTERING***______

```{r}
dataclust <- kmeans(model3fdata, centers = 3, iter.max = 100, nstart = 100)
```

Also, we need to determine and visualize optimal number of clusters.

Another method to ascertain the most favorable value of the K number of clusters is by using *Within Sum of Squares*, *Silhouette* and *gap_stat* plots. This advise us with 2 clusters.
To plot the 2 clusters, we can use *fviz_cluster()* function.

```{r}
fviz_nbclust(model3fdata, kmeans, method = "wss") 
fviz_nbclust(model3fdata, kmeans, method = "silhouette")
fviz_nbclust(model3fdata, kmeans, method = "gap_stat") 

dataclust <- kmeans(model3fdata, centers = 2, iter.max = 100, nstart = 100)
fviz_cluster(kmeans(model3fdata, centers = 2, iter.max = 100, nstart = 100), data = model3fdata)
```


The quality of the k-means partition is gauged by the **SSwithin**, and we want it to be as little as possible. Hence, we have 33.2%.
```{r}
dataclust$betweenss / dataclust$totss
```


Lastly, we can visualize clusters using the original variables where **x is Failure** and **y is Entropy_cooc.W.ADC**

```{r}
dataclust <- kmeans(model3fdata, centers = 3, iter.max = 100, nstart = 100)
model3data <- model3data |> mutate(cluster = dataclust$cluster)
model3data |> ggplot(aes(x = Failure, y = Entropy_cooc.W.ADC, col = as.factor(cluster))) + geom_point()
```


2.Hierarchical Clustering


In identifying the groupings in a data set, hierarchical clustering is a backup alternative to k-means clustering. Unlike k-means, in hierarchical clustering the number of clusters does not need to be preset, since this method can build hierarchy of clusters.


Also, before building a clustering model, a standardization of the data is prerequisite.
```{r}
fdata <- model3data%>%
  select_if(is.numeric) %>%  # select numeric columns
  select(-Failure.binary) %>%    # remove target column
  mutate_all(as.double) %>%  # coerce to double type
  scale()


```

# Hierarchical clustering using Complete Linkage

Resembling to k-means, we evaluate first the dissimilarity of observations using distance measures to get the agglomeration coefficient (AC). For us to sustain these values and specify the agglomeration method to be used either *"complete", "average", "single", or "ward.D2"*, is by using *hclust() function*.
```{r}
cdata <- dist(fdata, method = "euclidean") #dissimiliraty matrix

heirclustA <- hclust(cdata, method = "complete")
plot(heirclustA, cex = 0.6)
rect.hclust(heirclustA, k = 2, border = 1:4)
```

**AGNES**

We can also use agnes() function.
```{r}
set.seed(123)
heirclustB <- agnes(fdata, method = "complete")
heirclustB$ac
```

**DIANA**
```{r}
heirclustC <- diana(fdata)
heirclustC$dc

dianaA <- fviz_nbclust(fdata, FUN = hcut, method = "wss", 
                   k.max = 10) +
  ggtitle("(A) Elbow method")
dianaB <- fviz_nbclust(fdata, FUN = hcut, method = "silhouette", 
                   k.max = 10) +
  ggtitle("(B) Silhouette method")
dianaC <- fviz_nbclust(fdata, FUN = hcut, method = "gap_stat", 
                   k.max = 10) +
  ggtitle("(C) Gap statistic")
gridExtra::grid.arrange(dianaA, dianaB, dianaC, nrow = 1)
```

**Ward's method**
```{r}
heirclustD <- hclust(cdata, method = "ward.D2" )
resgroup <- cutree(heirclustD, k = 8)
table(resgroup)
```


3.Model Based

Next is model-based clustering. Unlike the previous clustering methods, model-based clustering automatically ascertains the favorable number of clusters. Also, the Gaussian mixture models is applied.


We may use the **Mclust() function*, which will leave *G = NULL* to necessitate the Mclust() to evaluate 1–9 clusters and select the optimal number of components based on BIC. 

```{r}
modelB1 <- Mclust(model3fdata[,1:10], G=3) 
summary(modelB1)
modelB2 = Mclust(model3fdata, 1:9) #from sir lecture 

#Error in plot.new() : figure margins too large
summary(modelB2)

```

# Plot results
```{r}
plot(modelB1, what = "density") #cannot plot 1:428 #dugay pag 1:10
plot(modelB1, what = "uncertainty")
```


```{r}
legargs <- list(x = "bottomright", ncol = 5)
plot(modelB1, what = 'BIC', legendArgs = legargs)
plot(modelB1, what = 'classification')
plot(modelB1, what = 'uncertainty')
```


```{r}
prbblts <- modelB1$z 
colnames(prbblts) <- paste0('C', 1:3)

prbblts <- prbblts %>%
  as.data.frame() %>%
  mutate(id = row_number()) %>%
  tidyr::gather(cluster, probability, -id)

ggplot(prbblts, aes(probability)) +
  geom_histogram() +
  facet_wrap(~ cluster, nrow = 2)
```


```{r}
uncertainty <- data.frame(
  id = 1:nrow(model3fdata),
  dataficlust = modelB1$classification,
  uncertainty = modelB1$uncertainty
)
uncertainty %>%
  group_by(dataficlust) %>%
  filter(uncertainty > 0.25) %>%
  ggplot(aes(uncertainty, reorder(id, uncertainty))) +
  geom_point() +
  facet_wrap(~ dataficlust, scales = 'free_y', nrow = 1)
```


```{r}
dataficlustB <- model3fdata %>%
  scale() %>%
  as.data.frame() %>%
  mutate(cluster = modelB1$classification) %>%
  filter(cluster == 2) %>%
  select(-cluster)

dataficlustB %>%
  tidyr::gather(product, std_count) %>%
  group_by(product) %>%
  summarize(avg = mean(std_count)) %>%
  ggplot(aes(avg, reorder(product, avg))) +
  geom_point() +
  labs(x = "Average standardized consumption", y = NULL)
```

**Conclusion**

By the results we can, therefore say that, using k-means clustering the best number of clusters is 2 with SSwithin = 33.2%. Also Hierarchical, gap statistics suggest 9 clusters with 84.90% AC and 84.29%. Lastly, model-based suggest 3 optimal number of clusters with BIC -2632.206.